\input a:scribe
\f:8=ff12
\def\FF{\f:8}
\majorhead{The tex Hackers Guide}

\head{Introduction}

This document is a guide to those tex users who are, for some
reason, dissatisfied with the way tex works. Since tex is a stage
one program (pasted up by some hacker with more guts than brains,
and it does what I want) it is to be expected that lots of people
will be dissatisfied with the way tex works. Therefore, this
document is a guide to adding things to tex allowing easy
integration into the tex distribution, and pointier to things
that I think need to be done.

The first section discusses extending the existing macro
packages.  The second section discusses adding new macro
packages, and how to document them with little pain. The final
section discusses changes to the actual tex code.

Before going into how these things can be done, there are two
things to point out. First, the tex package {\b is not in the
public domain!} I still own all copyrights, etc. I will,
however, allow redistribution of the package on three conditions:
{\verse
1) My copyright notices stay where they are.
2) The source is distributed with the package.
3) Fixes and improvements eventually come back to me.
}

Second, tex was written to satisfy me, and I am happy with it. It
was designed to let me use the power of {\FF Fancy Font} on my
math courses, and in writing letters.  It does an adequate job of
that. It probably won't do an adequate job of anything that
deviates too far from that.

\head{Changing the Macro Packages}

Probably the simplest thing in tex to change is the macro
packages themselves. The general rule is: {\b DON'T.} The macro
packages should be fairly complete. If you really want a large
set of new commands, the thing to do is to add a new macro
package that includes the commands you want. In this case, see
the next section.

\subhead{New Characters}

There are two exceptions to the general rule of not changing
macro packages. The first one is if you want to change the
characters set around. For example, you may decide you need a new
math mode character. Just print off a copy of the math fonts, and
add it - probably to \example{math2,} as \example{math1} is about
full. If you do this, send me a copy of the \example{edf} file
for the character, and the command name you used on the
character, and I'll see that the character/command pair winds up
in the next distribution. Maybe not in the same place, but it
will be there. I will also add that character to the
documentation.

Along these lines, there are two things that I would like to see
done, but don't have time to do. First, the Greek characters are
a bit bolder than I would desire, and not quite to the right
scale. This could be fixed.

Second, {\b all} the characters in the math2 font cause extra
lines to be added to the output. I don't know why these lines are
there, but it annoys me. If somebody (are you listening,
SoftCraft?) would fix this, I would greatly appreciate it.

\subhead{Changing Tex}

The second reason to want to change a macro package is to take
advantage of capabilities you have added to tex. There isn't much
I can say about this, except that you ought to add the new
commands to the documentation in the command list and the
appendices. Or, when you send them to me, I can do it, and see
that things get onto the tex distribution.

\head{Adding a New Macro Package}

If you don't like any of the macro packages currently available
for tex, you'll probably want to add your own. If you are trying
to make tex look like your favorite text formatter, I suggest
that you consider carefully which commands should and shouldn't
have arguments, based on the current 1K limit to argument sizes.
For example, when the \example{scribe} package was added - to
make tex behave like scribble - I made the \example{scribble}
inline commands take arguments, and those that affect the
environment were turned into simple commands.


You should also document your new macro package in a manner
similar to what is already there. The standard tex manual has
been designed so that any macro package documentation can be put
into it with very little work. Assume that your macro package is
called \example{gort.} Then you need to create the file
\example{mgort.tex} which documents the commands in the package.
This file should look like \example{mbasic.tex.} Place it on the
same disk as \example{mbasic.tex} (or whichever macro doc file
you use), and modify the \example{macros} command in the file
\example{texuman.tex} to use the name of the package in question.
This should do the trick.

You should also build two versions of the macro package. One
should include all the math mode commands, and use both math
fonts. The other should only include the miscellaneous math mode
commands, and not include the \example{math2} font. The reason for
this is that there are {\b lots} of math mode commands, and for
anything but mathematics, you won't need most of them. So the
second version of the macro package doesn't include them, and
also frees up another font file for the user. See the two basic
macro package files, \example{sbasic.tex} and \example{basic.tex,} for
examples.

I don't have any suggestions as to what kind of new macro
packages need to be built. Naturally - I built the ones I want.

\head{Changing tex Itself}

Finally, we come to the part that every true hacker probably went
to first. Well, it's not going to do them any good. I'm not going
to talk about the details of tex's internals. Instead, I will talk
about the major organization.

tex was designed to give me the ability to cleanly use the power
of {\FF Fancy Font} as quickly as possible: I was in the middle
of a very hairy algebra course when it was written, and wanted to
get tex up without destroying my grades. I almost succeeded.

\subhead{The Insides of tex}

Given this constraint, it should come as no suprise that the
basic algorithm that tex uses is stupid. To wit: tex reads in
characters until it finds a word break, and then outputs the word
and starts the next one with a space. If, at any time in this
process, a character should be read in that wouldn't fit on the
current line, tex starts a new line, and continues collecting
words. To compound tex's stupidity, I let \example{pfont} do all the
hard work: pagination, justification, etc. All tex does is word
filling, macro expansion, and environment nesting. Fortunately,
the result is very usable.

The last thing I have to say on the internals is that I tried
hard (well, reasonably hard) to make it easy for other people to
modify tex. Along these lines, all globals have nice, long names;
procedures tend to have descriptive names (\examplg{docom, setleft,}
etc), and every procedure is commented. Hopefully, you will have
little trouble figuring out what is going on internally.

\subhead{tex Wish List}

There are only three things that I can think of that I would like
to do to tex. I'm not going to do them - I'm going to play with
my now two-month old LISP processor.

First, modify tex so that it inputs an entire line and does its
own justification. This would allow the \example{rtmarg} command to be
turned on, and a couple of commands for handling quotes, etc
could be added to the \example{scribe} macro package.

Second, do the same thing as the first change, but vertically
instead of horizontally: have tex do its own pagination. This
would fix the \example{eject} kludge mentioned in the user manual,
allow footers and headers to be done reasonably, and is necessary
(but not sufficient) for adding footnotes to tex.

Finally, the major change. This is the change I was reserving for
version 2. Go look at the source code: you will observe that
the \example{processtext} function is the entry point for what amounts
to a recursive descent parser for tex files. To facilitate macros
and such things, the functions in this parser expect that one of
their arguments is a function that will return the next character
in the current input stream.

This parser needs to be changed so that its output can go
somewhere other than the output file, by giving all the
procedures a functional parameter that you pass characters to be
output to. Such a change would allow such things as footnotes and
endnotes. In addition, the input to the \example{ff} (or to the
header/footer commands, if they exist by then) could be run
through tex, so that you get to all the nice math mode characters.

\head{Last Note}

Well, so much for the things I think you should/can do to tex,
and what to do if you do them. I would like to remind all the
readers that if they are going to stay legal, they have to send
me any changes they implement. The appropriate addresses are
currently (Spring '83):
{\verbatim

Post Offal: Mike Meyer
P.O. Box 1749
Norman, OK 73070

Ma Bell: 405/360-2508

ARPANet: mwm@okc-unix

UUCPNet: decvax!duke!uok!uokvax!mwm

CNet: Ishtar - the Oklahoma CNode 406/364-1373
}

Obviously, I can't do a lot about it if you decide not to send me
updates - unless you start distributing tex yourself. I probably
won't get upset in that case, unless you are selling tex without my
name on it. Then I may get legally angry.

In any case, I hope you get as much good out of tex as I am.

{\closing
Happy hacking and may all your bugs be small,

Mike}
header/footer commands, if they exist by then) could be run
through tex, so th