% defines to make life easier for the macro package writer
\def\macros#1{\def\macname{#1}\def\fetchman{\input b:m#1}}

\macros{Basic}			% Which macro package are we documenting?
\input a:basic			% set up the environment
\ff{+nf +hl "tex USERS MANUAL\rDecember 1982"
+fl "Mike Meyer\rPage \s#"}
\f:9=ff12			% the fancy font fonts
\def\FF{\f:9}
\def\heading#1{\vskip 20pt{\bf #1}\vskip 10pt\par}
\def\subhead#1{\vskip 20pt{\it #1}\vskip 10pt\par}
\def\tex{{\rm T\FF E\rm X}}	% define some usefull strings
\def\metafont{{\ss METAFONT}}
\def\tm{{\FF TM}}
\def\cr{{\FF R}}
\def\pfont{{\ss pfont}}
\def\describe{\lftmarg +60\indent 0\indent -60}
\def\bold#1{{\bf #1 }}

\vskip 72 pt			% set up the title page
\ctrline{\bf SMALL tex USERS MANUAL}
\ctrline{\bf For the \macname\ Macro Package}
\vskip 24 pt
\ctrline{Mike Meyer}
\vskip 72 pt

This users manual for the small tex text formatting system is
intended for people familiar with some text formatter (not
necessarily \tex) and computers in general. The reader is assumed
to have used and to be familiar with the {\FF Fancy Font}
typesetting system.

\vskip 200 pt
\ctrline{copyright $\copyright$ 1982, Mike Meyer}
\eject
% now start the paper
\heading{Introduction}

Small tex (`tex' is never capitalized) is a micro-computer text
formatter based on Donald Knuth's \tex\tm\ computerized
typesetting system. tex (see?) shares with \tex\ user-definable
commands, multiple typefaces of various sizes and shapes, support
for typesetting mathematics,  and the ability to group text.
Unlike \tex, tex does not work with a character-definition
system, like \metafont, for output. Instead, tex uses {\FF Fancy
Font}\tm\ from SoftCraft for output. This package is not as
powerful as \metafont, and the missing abilities are reflected in
tex as missing features. Likewise, a z80 CP/M\tm\ system isn't a
DEC\tm\ 10, so many nice features of \tex\ that I considered
dispensable were dispensed with.

I wish to emphasize that tex is not meant for {\bf
all} micro-computer text formatting purposes. It is intended
specifically for formatting small documents that require more
abilities than most printers supply. Specifically, it was
designed to let me typeset papers for my courses in graduate
mathematics - which called for characters not generally available
on my printer. tex is also suitable for setting letters, etc.
However, for anything longer than about 20 pages, I don't recommend
it. I don't consider this an unreasonable restraint, as it takes
about 4 hours to print a 20 page document in {\ss +rd 0} mode.

\heading{What Does tex Do?}

tex provides a {\it mapping} from its input to input for \pfont\ 
that hopefully reflects the user's intent for the resulting
document's appearence. Unfortunately, tex is {\bf dumb.}

tex only knows about {\it words} and {\it lines.}
A {\it word} is anything that has whitespace on both sides {\bf in
the output.} Most especially, anything printed in math mode is
considered to be a single word. A {\it line} is a single line of
text in the output document.

As tex maps input to output, it reads in words, one
at a time, until the words won't fit on the output line it is
preparing. Then tex outputs the words that will fit on
the line, and starts the next line with whatever word wouldn't
fit on the current line.  This is insufficent since {\it
pages} are also important. Since tex ignores them, it can't do
anything about widowed lines - it doesn't know that they're
widowed. Likewise, you may occasionally tell tex to go to the
top of a page (with the {\ss eject} command) and tex, not
realizing that it is already at the top of a page, will leave a
blank page behind when it obeys your command.

This behavior can be modified by magic characters in the input
stream. tex considers the same characters magic that \tex\ 
considers magic in its {\ss basic} package: `\\', `\%', `$\$$',
`\{', and `\}'. These have the same meanings as the do in \tex,
but cannot be changed as they can in \tex. They can be added to
the output stream in the same manner as they are in \tex: by
preceding them with a `\\' character, which effectively makes
them commands.

The `\\' character is used to indicate to tex that a command
follows. Commands are {\it action} words - they tell tex to do
something, usually immediately.

The `\%' character signals a comment. When tex sees a `\%' in the
input stream, it ignores any preceeding whitespace, all
characters to the end of the line, and any whitespace following
the end of the line. {\it Whitespace,} if you were wondering,
consists of blanks, tabs, and new line characters - things that
don't leave ink on the paper.

The $`\$'$ character indicates a change of mode - to or from one
of the two math modes. $`\$''$s are used in pairs, either as:
\ctrline{\ss$\$$text for math mode$\$$} or:
\ctrline{\ss$\$\$$text for display mode$\$\$$} In math mode and
display mode above, tex recognizes extra commands, and ignores
spaces (except as needed to separate out commands).

Finally, the `\{' and `\}' characters group text and arguments to
commands. When freestanding, they indicate that commands that
alter global properties of the text should apply {\bf only} to
the text contained in them. These commands will be noted as we
encounter them. For commands that require textual arguments, `\{'
and `\}' set off arguments that are longer than one character.

The following sections will discuss tex input syntax, the builtin
commands for tex, and the \macname\ macro package in particular.
Examples will be drawn from the input used for the manual. In
particular, references to "line <number> of the input" refer to
the appropriate line number in the file {\ss texuman.tex.} Also,
there are occasional references to things done nearby in the
text. These also indicate that you should check the appropriate
input file. Therefore, if you have not printed off copies of the
tex source for this manual (both {\ss texuman.tex} and
{\ss\macname.tex}) you should do so now. I recommend {\ss pip
prn:=} to get numbered listings.

\heading{What the Magic Characters Do}

As previously mentioned, there are several magic characters that
change the way tex behaves. These are `\\', `\%', `$\$$', and the
curly braces `\{' and `\}'. This section discusses their
behavior and use in some detail.

\subhead{The Command Character and Command Formats}

The magic character `\\' introduces a tex command. Commands to
tex have several formats. The basic format for all commands is a
`\\' character, followed by the name of the command, possibly
followed by arguments. In any of the formats, any following
whitespace is ignored by tex. Therefore, if you want the output
from a command followed by a space, you should probably issue a
`hard space' command. This is usually {\ss`\\\ '.} If you are not
sure, you can always force tex to output a space by using the
string {\ss \{\}} followed by a space.

Command names come in two forms. The first consists of a single
nonalphabetic character. Examples would be the commands used to
get magic characters into the text. Around line 190 in the input
is a good example of this type of thing. The second form consists
of a string of alphabetic characters terminated by some
nonalphabetic character. Usually, this nonalphabetic character is
a blank. In this form, the case of the first character is
significant, but the case of any following characters is ignored.
For example, {\ss \\sigma, \\sIGMA,} and {\ss \\sIGma,} are all
the same command, and all different from {\ss \\Sigma.}

Commands usually have either one or zero arguments. Examples of
the first case include commands to insert character strings into
the text stream, as in the {\ss tex} command used on line 26 of
the manual, in the abstract. The other use for commands without
arguments is to change some global parameter of tex, say the
typeface being used to print characters. The {\ss FF} command on
line 12 of the input is a good example of that.

If a command has a single argument, this command is known as a
{\it token.} Tokens come in two forms. The general case is that
they are a sequence of text between `\{' and `\}.' In such cases,
{\bf the argument is limited to 1024 characters.} An example of
this form is the {\ss macros} command on line 4 of the document.
The argument to this command is the string {\ss \macname.} The
other form for a token is a single, non-whitespace character.
Generally, these are used after the non-alphabetic commands, as
in {\ss \\+x,} which would be the command named {\ss +} with an
argument of {\ss x.}

Finally, there are the commands which have a strange argument
syntax. These either require more information, or are done in
some specific manner to be compatible with \tex. These commands
are {\ss def, vskip, hskip, f:, input} and {\ss math:.}

\subhead{Comments}

Comments in tex are introduced by the character `\%', and
continue from the \%\ to the end of input text line. When a bare
\%\ is seen, tex will ignore any white space preceding the \%,
any text between the \%\ and the end of the line, and any
whitespace following the comment. For examples of comments, see
the first 20 or so lines of the input for this document. To get a
\%\ into the output text, use the {\ss \\\%} command.

\subhead{The Math Modes}

\tex\ was originally designed for typesetting text with a high
content of mathematics. tex has the same intent, and uses the
\tex\ techniques for specifying equations. Specifically, an
equation is introduced by either `$\$$' or `$\$\$$'. A single
$\$$ introduces an inline equation, like $a=3x+b.$ The
double $\$$, $\$\$$, introduces an equation centered on
a line by itself, as: $$a=3x+b$$

While in math mode, tex changes to an italic-like font for normal
characters. Any whitespace in an equation in math mode is
ignored, except for the space command `\\\ '. The advantage of
math mode is that a {\bf large} number of commands become
available to output symbols not normally available.  A complete
list of these commands can be found in the appendix. Those of
general interest are printed here, as pairs of {\it name, graphic
representation}:{\fill 0\ss{}

\lft{$\$$, $\$$}	\ctr{$\#$, $\#$}	\rt{dag, $\dag$}
\lft{ddag, $\ddag$}	\ctr{ldots, $\ldots$}	\rt{ldotss, $\ldotss$}
\lft{circ, $\circ$}	\ctr{bullet, $\bullet$}	\rt{copyright, $\copyright$}
\lft{section, $\section$}}

In addition, the two math mode commands {\ss +, -} generate
superscripts and subscripts, and are occasionally useful outside
of equations.

As in \tex, there are lots of techniques for making the equations
from tex look better than just typing commands in math mode.
These will be discused in the next version of this manual.

\subhead{Nesting Text and Tokens}

The characters `\{' and `\}' are used in tex to group arguments
to commands, and to indicate {\it environment nesting.} Before I
can discuss the brace characters, I need to discuss environments,
and the nesting of environments.

It must be understand that the text formatting done by tex does
not take place in a vacuum. The output of tex is governed by
several knobs of tex that the user can set. These knobs comprise
such things as the current typeface, the margins, etc. The
current {\it environment} is the current setting of all of tex's
knobs.

The {\it nesting} of one environment inside of another implies
that the environment is changed temporarily for a piece of text,
and after the text is processed, the environment is restored to
its previous state. This concept will be familiar to anyone wh
has worked with an ALGOL-like language, and indeed works in a
similar manner. You can nest an environment inside of an
environment that is already nested. In fact, you can nest
environments in this manner arbitrarily deep (until you run out
of stack space, anyway). This facility is very powerful, and
makes many things easy that would be difficult in a text
formatter such as roff.

In tex, environments are nested with the pair of characters `\{'
and `\}.'  When not used to delimit arguments, the character `\{'
indicates that a new level of environment nesting is being
entered. Commands that change the current formatting environment
may be issued, and will affect only the text enclosed between the
open and close braces. To see how this nesting works, consider the
subheadings in the next section. These nest an environment using
an italic typeface for the main text of the subheading, and nest
another environment inside of that one to use a sans serif
typeface for the command names.

{\lftmarg +20\indent 0\indent -20\sc The true value of this
technique may be seen in this paragraph, which change the margin
and indentation to become an outdented paragraph, and changes the
typeface to script. Then, when the paragraph is finished, the
environment closed, and things will return to normal, regardless
of what `normal' was before.}

Likewise, environment nesting can be used in letters. For
example, a typical heading might look something like:

{\fill 0{\lftmarg 450{}
P.O. Box 1749
Norman, OK 73070
Jan. 20, 1982}

Hank Harlison
Dr. Mobs Urinal
Somewhere, CA xyzzy
}


Find the commands in the source file that were used to produce
this text to see how easy this is for tex, and then compare it to
what would have to be done in a roff-like formatter.

The other use for the brace characters is to delimit an argument
to a command. This is discussed fully in the subsection on
command characters and formats. The important thing to note at
this point is that the braces used to delimit an argument {\bf do
not cause environment nesting}. In some cases, the command can be
built to cause this to happen.  None of tex's builtin commands do
this.

\heading{The Builtin Commands to tex}

Following is a list of the commands that tex will always know
about, no matter what macro package you are using. This list is
not a complete list of all the commands a normal tex user will use,
but merely covers those that are useful for setting
up a document. There are other builtin commands that will
probably be useful if you are building or modifying a macro
package. See the Appendices for a complete list of builtin
commands. The majority of the commands normally used will
probably be defined in a macro package. See the following
section for a list of these commands.

\subhead{The {\ss def} Command}

The single most useful builtin command is {\ss def.} It allows
the user to define new commands that look, and act, just like
builtin tex commands. Defined commands can have either alphabetic
or non-alphabetic names, and can have either of the two standard
argument forms - none, or a single token.

Examples of the {\ss def} command can be found on lines 9 through
15. Only lines 10 and 11 define commands with arguments. The
other lines all demonstrate the standard, non-argument {\ss def}
command format:
\ctrline{\ss \\def\\name\{text of define\}}

This string in the input text causes the command {\ss name} to be
defined, such that every time the string {\ss \\name} appears in
the input stream, {\ss text of define} will appear in the output
stream. For example, the defintion for {\ss tex}  appears on
line 13. This line defines the {\ss tex} command such that
every time the string {\ss \\tex} appears in the input, the
string {\ss \{\\rm T\\FF E\\rm X\}} will be fed to tex, causing
the symbol \tex\ to be printed on the page.

The second form of the {\ss def} command is shown on lines 10 and
11 - a {\ss def} command with an argument. The general form of
this is:
\ctrline{\\def\\name#1\{text with possible #1 expansion\}}

This form of the {\ss def} command works the same as the
argumentless form of {\ss def,} except that, when the text for
the definition is being used by tex, any occurence of the string
{\ss #1} will be replaced by the argument. For example the {\ss
heading} command, which is defined on line 10, when used on line
33 will expand to the string {\ss\{\\vskip 20pt\{\\bf
Introduction\}\\vskip 10pt\\par\}.}

A particularly good example of the power of the {\ss def} with a
single argument can be found on line 2 of the input. This line
defines the {\ss macros} command, which has one argument.
Immediately below this, on line 4, the {\ss macros} command is
used, with the argument of {\ss \macname.} This causes the text
of the {\ss macros} command to be read by tex, which defines two
more commands - the {\ss macname} command, and the {\ss fetchman}
command. The {\ss macname} command just puts the argument to the
{\ss macros} command in the input stream. The {\ss macname}
command is used in the preceding sentence to get the argument of
the {\ss macros} command on line 4 into the text, and on line 19
to get the name of the macro package this manual documents into
the subtitle. The {\ss fetchman} command tells tex to input the
parts of the manual that are specific to some macro package, and
is used to do so later on in this document. Note that the {\ss
macros} command demonstrates that you can use the {\ss #1}
sequence in a definition expansion as often as needed. It is left
as an exercise to the reader to determine if a definition can
define commands with arguments, and how this would be done if it
can be done.

\subhead{Positioning Commands - {\ss vskip, hskip, ctr, rt, lft, eject}}

There are five commands built in to tex for positioning text on
the page. These are {\ss hskip, vskip, ctr, rt, lft} and {\ss
eject.} The only two that will generally be used are {\ss vskip}
and {\ss hskip.} These both have the same format: \ctrline{\ss
\\hskip xx pt} or \ctrline{\ss \\vskip yy pt}

{\ss hskip} causes a {\bf h}orizontal {\bf skip} in the output
text of xx {\it pts.} A {\it pt} (for hskip) is one 120$\+{th}$ of
an inch. For example, the following text is produced by the
input sequence {\ss |\\hskip 50 pt|}:|\hskip 50 pt|

{\ss vksip} causes a {\bf v}ertical {\bf skip} in the output text
of yy pts. For {\ss vskip,} a pt is {\bf one 72$\+{nd}$} of an
inch. {\bf NOTE THE \underline{DIFFERENCE}} between the unit for
{\ss vskip} and {\ss hskip.} This is unfortunate, but it is also
unavoidable, being imposed by {\FF Fancy Font}. An example of the
use of {\ss vskip} is in the definition for {\ss heading} and
{\ss subheading,} where {\ss vskip} is used to get appropriate
amounts of vertical spacing around the headings.

In both {\ss vskip} and {\ss hskip,} the string {\ss pt} is {\bf
required.} If it is left off, tex will complain, and eat the
character following the {\ss v[h]skip} command.

The {\ss eject} command is used to get to a new page. An example
of this is seen on line 31, where it is used to get to a new page
to start the manual after the title page has been printed. As has
been pointed out before, tex does not know about pages, so {\ss
eject} issued at the top of a page will cause a clean sheet of
paper to be printed. For this reason, it is inadvisable to use
lots of ejects or, if you do use lots of ejects, to number
your pages.

The three commands {\ss lft, rt} and {\ss ctr} are relatively
dangerous, and should only be used by experienced tex users:
those who know what the commands actually do. In general, the only
place they will be used is in producing `tables,' or the closest
approximation that tex can make to tables. For an example of this,
see the section around line 200 in the input document on math
mode.

The only really useful feature any of these offer is centering
lines of text. Check the definitions provided by the macro
package you are using for a specific centering command.

\subhead{Fill and Nofill}

One of the nonstandard form commands is the {\ss fill} command.
Fill affects the behavior of tex in a major way. The general form
is {\ss \\fill x,} where x is either zero or one. If x is one,
then tex behaves as previously described, treating newline
characters as spaces, and packing as many words per line as
possible. If x is zero, then tex {\bf stops doing word filling.}
This means that every newline character in the input causes one
newline in the output. Of course, if a line would be to long to
fit on the page, tex will break it if you are not in fill mode.
Additionally, tex does not justify the right margin when in it is
not in fill mode.

{\fill 0
For example, this text is not
filled in. You can see the difference
it makes in the output.}

{\ss fill} is useful for such things as addresses on letters,
tables, etc. Note that fill is one of the global formatting
parameters for tex, so it is restored to its old value when
leaving a nested environment. The tables in the discussion of
math mode demonstrate this.

\subhead{Paragraph Commands - {\ss par, indent, lftmarg}}

In general, tex does not know about paragraphs. However, the
three commands {\ss indent, lftmarg} and {\ss par} affect the shape
of the document by affecting the way paragraphs are laid out on
the page.\par The simplest of these is {\ss par.} {\ss par} is
causes tex to start a new paragraph {\bf immediately.} For
example, this paragraph was started with a {\ss par} command in
the input stream. Note that if you are in fill mode, two
consecutive newline characters are synonymous with the {\ss par}
command. This is usually preferable, since most text editors that
know about paragraphs will recognize that sequence for a
paragraph. Additionally, it visually breaks out paragraphs in the
input.

{\ss lftmarg} and {\ss indent} are used to change the shape of a
paragraph. {\ss lftmarg} is used to set the left margin of the page,
and {\indent} is used to set the amount of paragraph indentation.
The form for both {\ss lftmarg} and {\ss indent} is the same:
\ctrline{\ss \\lftmarg sxx} or \ctrline{\ss \\indent sxx} In these
forms, the {\ss xx} is some number of pts as defined for {\ss hskip},
one 120$\+{th}$ of an inch. The {\ss s} in the argument is used to
allow finer control. If it is a blank, the value of {\ss lftmarg} or
{\indent} is set to the numeric value of xx. If s is a `+' (or
`-'), then the value of xx is added (or subtracted) from the
current value of {\ss lftmarg} or {\ss indent.}

Since both {\ss lftmarg} and {\ss indent} are part of texs
formatting environment, these values will nest with curly braces,
so some rather nice effects can be achieved. For example, a
quotation can be done by setting up a block of text in curly
braces that start with {\ss \\lftmarg +20.} This will indent the
paragraph(s) in the quotation by 1/6$\+{th}$ of an inch.
Similarly, outdented paragraphs can be set up by changing the
left margin to an appropriate value, then setting indent to 0,
and finally subtracting whatever you deem appropriate from
indent. See the appendices for examples of this.

\subhead{Miscellaneous Commands}

Finally, there are four commands varying in usefulness from
considerable to nearly none. These don't fit in any of the above
sections, as each is unique.

The most useful of these three commands by far is the {\ss input}
command. You will use it in almost every document you use tex on,
generally to fetch the macro package you need for that document.
The syntax of the {\ss input} command is: \ctrline{\\input filename}

This causes tex to go look for the file {\ss filename.tex}, and
start reading input from that file. This input is inserted
directly into the stream, in the middle of the word if that's where
the {\ss input} command occurred. Inputting text via the input
command is {\bf not} like nesting text inside a pair of curly
braces - any changes made to the global environment in the input
file will affect text following the {\ss input} command. For
examples, see lines 2 and 5 in the input text.

The second command is the {\ss underline} command. This command
has a single token as an argument, as in {\ss
\\underline\{text\}.} This will cause {\ss text} to be
underscored. {\ss underline} obeys all the constraints placed on
normal, one-argument commands.

The third command is the {\ss ff} command. This command passes a
single argument to \pfont\ when it is invoked by tex. Lines 6 and
7 in the input show how I used this to put headers and footers on
the manual. Note that only the last occurrence of this command in
the text is actually used.

Finally, there is the {\ss f:} command. This command is used to
declare a {\FF Fancy Font} font file to be associated with a font
number, and to change to that font during the input stream.
Normally, there are only two font numbers you can use, 8 and 9.
The syntax to associate a font file with font number 8 is: {\ss
\\f:8=file}. To cause a change to font number 8, the command {\ss
\\f:8} is issued. Normally, you won't use this version of the
command, as you (or whoever wrote the macro package you are
using) will have defined a command to change fonts for you. Lines
8 and 9 in the input file show font 9 being associated with the
file ff12.fon, and the the command {\ss FF} being defined to
change to that font. From then on, when I want to use that font
(like I did in this paragraph), I issue the {\ss FF} command.

\heading{Making use of the \macname\ Macro Package}

\fetchman

\heading{Caveats}

This manual is not the last word on using tex. It is meant to be
enough of an introduction to let new users start with tex, and
get some work done. It does not document all the features and
behavior of tex: the source code does that. If you don't have the
source, somebody is distributing copies of tex in violation of
the release agreement. Get in contact with me in that case.

If you have problems, or need help with tex, get in contact with
me. The appropriate addresses are currently:

{\fill 0
U.S. Post Offal:{\lftmarg +50
Mike Meyer
P.O. Box 1749
Norman, OK 73070}

Bell Telephone:{\lftmarg +50
405/360-2508}

ARPANet:{\lftmarg +50
mwm@okc-unix}

UUCPNet:{\lftmarg +50
decvax!duke!uok!mwm}

CNet:{\lftmarg +50
The Oklahoma Cnode - Ishtar: 405/329-1373}}


\heading{Appendix A}

This appendix is the most important single part of tex users
manual.  It lists all the error messages that tex issues. The
format of an error message is {\ss tex: text of error message.}
The following list contains the text of the error message,
followed by comments about the probable cause and possible fixes.

{\describe

\bold{Invalid file name:} You told tex to format a file with a
name that is longer than is valid for the operating system. File
names used with the {\ss input} command cannot contain `.'
character. If you used a `.' in such a file name, the {\ss
Invalid file name} error message can be issued. In either case,
check the filename on the command line, or in the appropriate
{\ss input} command.

\bold{Can't open file:} tex couldn't find the input file. Check
the spelling of the filename on the command line or in the
appropriate {\ss input} command.

\bold{Can't create file:} tex couldn't create the output file.
Possibly your disk is full.

\bold{Missing right brace in file:} There is a missing right
brace in the named file. Verify that every open brace has a
matching close brace.

\bold{Document completed in math mode} tex found the end of the
file while it was in math mode. You undoubtedly forgot to put the
final $\$$'s on math mode or display mode text. Use the {\FF Fancy
Font} {\ss +sd} mode to find where tex stops outputting spaces.
This is probably where the missing $\$$ belongs.

\bold{Negative left margin, left margin set to 0} You tried to
change the left margin to a negative value. It was set to zero.
You probably tried a {\ss lftmarg} command with a decrement
argument that was larger than the current left margin.

\bold{Indent to small, set to:} You tried to change the indent to
a value that, when added to the left margin, created a negative
offset. tex set the indent to the negative of the left margin,
which is probably close to what you wanted (pray, pray). Look for
sequences like {\ss \\indent 0\\indent -x} where x is larger than
the value of the left margin.

\bold{Can't create temp file} tex couldn't create the file it
uses for passing parameters to \pfont. Possibly your disk is
full.

\bold{Write error on file:} The write to the file tex uses for
passing parameters to \pfont\ failed. Possibly your disk is
full.

\bold{Can't execute pfont} tex couldn't find the \pfont\ command.
Check to see that it is on the correct disk.

\bold{Out of buffer space} tex ran out of buffer space for
input. This is highly unlikely. The only thing I can think of
that would cause it is having the margins set to allow a large
number of characters per line (either wide margins or small
characters, or both), and being in math mode.

\bold{Internal error in mathmode: inchar =} tex found itself in
display mode while not in math mode. Write down the entire output
line and get a hacker to look at your input and the version of
tex you are running.

\bold{Leaving display mode with $`\$'$} You closed display mode
text with a single `$\$$.' Either you closed display mode
incorrectly, or you forgot to close some math mode text. In the
latter case, you should also get {\ss Math only command$\ldotss$}
messages.

\bold{Leaving math mode with a $`\$\$'$} This is the converse of
the previous error message. The reasons and fixes are the same.

\bold{Domath compiler error!} Something that shouldn't be
possible happened. Write down the error message, and get a hacker
to look at the your tex input, the source for the version of tex
you are running, {\bf and} the version of the compiler that was
used to compile it.

\bold{Illegal command:} tex found a command that it didn't
recognize. You probably made a typo in the command name. Check
your spelling.

\bold{Illegal font command:} There is a command that looks like
a font command ({\ss \\f} followed by a non-alphabetic character)
that is not a valid font command. Normally, font commands are set
up in the macro package, so this shouldn't occur for the normal
user. If it does, check the macro package you are using.

\bold{Illegal font character:} A font or math command occured
in the text that changed to an illegal font. See the previous
error meesage for discussion.

\bold{Can't open font file:} Somebody is trying to use a font
file that doesn't exist. Once more, check the macro package you
are using. Also verify that all the font files it uses are where
it expects them.

\bold{Bad font file:} The named font file contains data that tex
doesn't understand. Possibly it or tex is corrupt. Get a fresh
copy of the files. If that fails, verify that the version of tex
you are using is compatible with the \pfont\ you are using.

\bold{Invalid character:} You tried to use the builtin {\ss
char} to print a non-ascii character. If you don't use the {\ss
char} command, check the macro package you are using.

\bold{Math only command issued outside math mode} The math mode
commands complain about this if they are used outside of math
mode. In most cases, the command will do what you wanted, but not
fixing the cause of the problem is bad form. If you get many of
these, you probably forgot to close some math mode text properly.

\bold{Bad define character.} A {\ss def} command was encountered
that didn't end in a `\\'. Find the appropriate define, and fix it.

\bold{Define too long:} You tried to define a command whose name
was longer than tex can accept. Change the name of the defined
command.

\bold{Command too long:} You issued a command that was too long.
This is an illegal command, as you couldn't have defined it in
the first place. Find the occurence of the substring that tex
printed, and change it to a legal command.

\bold{Bad define:} The name of a defined command was not
followed by either a `\{' or a `$\#.$' Find the appropriate {\ss
def} command and fix it.

\bold{bad define parameter.} You tried to use a parameter number
other than 1 in a define. Currently, tex only allows one
parameter on a command. Change the appropriate define.

\bold{Too many defines:} The number of allowable defines in tex
has been exceeded. The named command was not defined. The fix is
either to use fewer defines, or reconfigure tex to allow more
defines.

\bold{Out of string space} You have run out of memory. Either
get more memory, decrease the size of defined commands, or
reconfigure tex.

\bold{Illegal command \\math} The builtin command {\ss math} has
been misused. Change the usage to correspond to {\ss \\math=x,}
where {\ss x} is an appropriate font number.

\bold{Need `pt' to follow hskip number} You left off the {\ss
pt} from an hskip command. tex will eat the charcter following
the misformed hskip command. Fix the offending hskip.

\bold{Need `pt' to follow vksip number} Identical to the
previous error, except that it applies to vskip instead of hskip.

\bold{Input token too long} An argument to a command exceeded the
maximaum length. Check commands with long arguments, and break
the offending command up into two separate commands.

}		% end of the appendix A

\heading{Appendix B}

This appendix describes the builtin commands of tex. For each
command, its form and usage is is discussed, as well as its
exact behavior in terms of the \pfont\ commands that it issues.
This appendix will probably be of use only to those building
macro packages. Everybody should be aware of the names of the all
the builtin commands, as you cannot use these names for defined
commands.

{\describe

\bold{f} The {\ss f} command is used to change fonts, or to
associate a font file with a font number. The font numbers
currently valid are 0 through 9, and attempts to use any others
will (hopefully) result in error messages. The form used to
associate a font file with a font number is {\ss \\f:x=name,}
which associates font number {\ss x} with font file {\ss
name.fon.} This causes no \pfont\ commands to be issued, but does
cause tex to load the named font file. The form used to change to
a font number is {\ss \\f:x}, which changes to font {\ss x.} This
causes the \pfont\ command {\ss \\fx} to be issued.

\bold{math} The {\ss math} command is used to tell tex which font
is to be use for math mode. The format is {\ss \\math:x,}  which
causes font number {\ss x} to be used for math mode text.

\bold{def} This defines a command, and its use is discussed in
the main body of the manual. It causes no \pfont\ commands to be
issued.

\bold{fill} The {\ss fill} is used to turn fill mode on and off.
Its use is discussed in the main body of the text. The current
fill mode is part of the tex environment. If used to turn fill
and justification on, a {\ss \\j} command is issued to \pfont.
Otherwise, a {\ss \\k} command is issued to \pfont.

\bold{lftmarg} This command is used to change the left margin. Its
use is discussed in the main body of the text. The left margin is
part of the tex environment. This command causes no \pfont\ commands to be
issued.

\bold{indent} The {\ss indent} command is used to change the
paragraph indentation tex is using. This indentation is part of
the tex environment. The use of the {\ss indent} command is
discussed in the main body of the manual. It causes no \pfont\ 
commands to be issued.

\bold{rtmarg} The {\ss right} command is analogous to the {\ss
lftmarg} command. It is currently disabled.

\bold{mathonly} This command checks to see that tex is currently
in math mode, and issues an error message if not. The correct
usage is merely {\ss \\mathmode.} This command should be used by
builders of macro packages to flag commands that should only be
issued in math mode. No \pfont\ commands are issued.

\bold{par} The {\ss par} command forces a paragraph output. It
could probably be made a define, but I don't see any point in it
yet. Its usage is discussed in the main body of the manual. No
\pfont\ commands are issued.

\bold{eject} The {\ss eject} command is used to force a new page.
The string  {\ss \\b<newline>\\p} is issued to pfont. Therefore,
the trailing line will not be justified, and the next line will
start on a new page.

\bold{ctr} This command is used to center text. The form is {\ss
\\ctr\{text\}.} The \pfont\ command {\ss \\c} is issued, followed
by the text of the argument. Note that \pfont\ will center any
text that follows the {\ss ctr} command, unless other commands
are issued.

\bold{rt} The {\ss rt} command is used to force text to the right
side of the line. Usage is {\ss \\rt\{text\}.} The pfont command
{\ss \\r} is isued, followed by the text of the argument,
followed by a newline character.

\bold{lft} The {\ss lft} command is used to force text to the
left side of the line. Usage is {\ss \\lft\{text\}.} The pfont
command {\ss \\b} is issued, followed by the text of the
argument. Hence, the total effect of the {\ss lft} command is to
turn off justification for the line it's issued on.

\bold{vskip} This command is used to force a vertical skip down
the page. Usage is discussed in the main body of the manual. The
\pfont\ command {\ss \\b} is issued, followed by a {\ss \\v}
command to cause the skip. Thus, the line the {\ss vskip} is
issued on will not be justified, and the next line will occur the
appropriate number of points down the page.

\bold{hskip} The {\ss hskip command} is used to skip horizontally
across the page. Usage is discussed in the main body of the
manual. The \pfont\ command {\ss \\i} is issued to cause the skip.

\bold{char} This command is used to force some non-printable
ascii character into the output stream. Usage is {\ss \\char
xxx}. The ascii character corresponding to the decimal value of
{\ss xxx} will be inserted into the output stream. Normally, the
appropriate ascii character is output. However, there are a
couple of cases where the \pfont\ command {\ss \\d} is used.

\bold{underline} The {\ss underline} command is used to underline
text in the output document. Usage is {\ss \\underline\{text\}.}
The output to pfont is {\ss \\utext of argument\\u.}

\bold{input} The input command is used to fetch text from another
file. Usage is discussed in the main body of the manual. No
\pfont\ commands are generated.

\bold{ff} This command is used to pass page formatting
information to \pfont. Usage is discussed in the main body of the
manual. No pfont commands are generated, but its argument is
passed to pfont on the command line.

}		% end of Appendix B

\heading{Appendix C}

This appendix  describes the commands available in math mode.
These commands should be independent of the macro package you are
using. The command set is broken up into pieces. The description
of each piece of the command set consists of a brief discussion
of that particular piece, followed by a table that has each
command listed beside the resulting graphics character.

The first piece of the command set is the Greek alphabet. The
upper case Greek characters that are not also part of the English
alphabet are included, as:{\fill 0\ss{}

\lft{Gamma, $\Gamma$}	\ctr{Delta, $\Delta$}	\rt{Xi, $\Xi$}
\lft{Theta, $\Theta$}	\ctr{Lambda, $\Lambda$}	\rt{Pi, $\Pi$}
\lft{Phi, $\Phi$}	\ctr{Sigma, $\Sigma$}	\rt{Upsilon, $\Upsilon$}
\lft{Omega, $\Omega$}	\ctr{Psi, $\Psi$}}

The entire lower case Greek alphabet is available, including both
forms of theta and and phi. Note that the {\bf case of the first
character} in the command is significant, and that the case of
the first character is the only thing that differentiates the
{\ss gamma ($\gamma$)} from the {\ss Gamma ($\Gamma$)} command.
The commands for the lower case Greek characters are:{\fill 0
\ss{}

\lft{alpha, $\alpha$}	\ctr{beta, $\beta$}	\rt{gamma, $\gamma$}
\lft{delta, $\delta$}	\ctr{epsilon, $\epsilon$}\rt{xi, $\xi$}
\lft{theta, $\theta$}	\ctr{eta, $\eta$}	\rt{iota, $\iota$}
\lft{kappa, $\kappa$}	\ctr{lambda, $\lambda$}	\rt{mu, $\mu$}
\lft{nu, $\nu$}		\ctr{pi, $\pi$}		\rt{phi, $\phi$}
\lft{rho, $\rho$}	\ctr{sigma, $\sigma$}	\rt{tau, $\tau$}
\lft{upsilon, $\upsilon$}\ctr{omega, $\omega$}	\rt{chi, $\chi$}
\lft{psi, $\psi$}	\ctr{zeta, $\zeta$}	\rt{vartheta, $\vartheta$}
\lft{varphi, $\varphi$}}

The upper case script characters are also available in math mode.
These characters are from the standard {\FF Fancy Font} script
character set, which is probably available with the macro package
you are using. They are provided in math mode as a convenience
for the user. They are:{\fill 0\ss{}

\lft{Ascr, $\Ascr$}	\ctr{Bscr, $\Bscr$}	\rt{Cscr, $\Cscr$}
\lft{Dscr, $\Dscr$}	\ctr{Escr, $\Escr$}	\rt{Fscr, $\Fscr$}
\lft{Gscr, $\Gscr$}	\ctr{Hscr, $\Hscr$}	\rt{Iscr, $\Iscr$}
\lft{Jscr, $\Jscr$}	\ctr{Kscr, $\Kscr$}	\rt{Lscr, $\Lscr$}
\lft{Mscr, $\Mscr$}	\ctr{Nscr, $\Nscr$}	\rt{Oscr, $\Oscr$}
\lft{Pscr, $\Pscr$}	\ctr{Qscr, $\Qscr$}	\rt{Rscr, $\Rscr$}
\lft{Sscr, $\Sscr$}	\ctr{Tscr, $\Tscr$}	\rt{Uscr, $\Uscr$}
\lft{Vscr, $\Vscr$}	\ctr{Wscr, $\Wscr$}	\rt{Xscr, $\Xscr$}
\lft{Yscr, $\Yscr$}	\ctr{Zscr, $\Zscr$}}

There is a large selection of binary relations. These are:{\fill 0\ss{}

\lft{normal, $\normal$}	\ctr{noteq, $\noteq$}	\rt{eqv, $\eqv$}
\lft{supset, $\supset$}	\ctr{subset, $\subset$}	\rt{in, $\in$}
\lft{notin, $\notin$}	\ctr{perp, $\perp$}	\rt{le, $\le$}
\lft{ge, $\ge$}		\ctr{onlyif, $\onlyif$}	\rt{implies, $\implies$}
\lft{if, $\if$}		\ctr{iff, $\iff$}	\rt{mapsto, $\mapsto$}
\lft{lsls, $\lsls$}	\ctr{grgr, $\grgr$}	\rt{simeq, $\simeq$}
\lft{approx, $\approx$}	\ctr{doteq, $\doteq$}}

There is also a variety of binary operators available in tex. These
are:{\fill 0\ss{}

\lft{pm, $\pm$}		\ctr{mp, $\mp$}		\rt{times, $\times$}
\lft{div, $\div$}	\ctr{odot, $\odot$}	\rt{oplus, $\oplus$}
\lft{odiv, $\odiv$}	\ctr{ominus, $\ominus$}	\rt{otimes, $\otimes$}
\lft{cdot, $\cdot$}	\ctr{union, $\union$}	\rt{inter, $\inter$}
\lft{and, $\and$}	\ctr{or, $\or$}}

In addition to the binary operators, the following unary
operators are avialable:{\fill 0\ss{}

\lft{partial, $\partial$}\ctr{surd, $\surd$}	\rt{int, $\int$}
\lft{oint, $\oint$}	\ctr{sum, $\sum$}	\rt{prod, $\prod$}
\lft{prime, $\prime$}}

There are also a several different flavors of arrows. Note that
there is no right arrow as such. The arrows that are there
are:{\fill 0\ss{}

\lft{up, $\up$}		\rt{bigup, $\bigup$}
\lft{down, $\down$}	\rt{bigdown, $\bigdown$}
\lft{left, $\left$}	\rt{bigleft, $\bigleft$}
\lft{right,$\right$}	\rt{bigright, $\bigright$}}

The following miscellaneous math symbols can also be used:{\fill 0\ss{}

\lft{thereis, $\thereis$}\ctr{forall, $\forall$}\rt{therfor, $\therfor$}
\lft{cdots, $\cdots$}	\ctr{cdotss, $\cdotss$}	\rt{nabla, $\nabla$}
\lft{imag, $\imag$}	\ctr{natural, $\natural$}\rt{rat, $\rat$}
\lft{real, $\real$}	\ctr{integer, $\integer$}\rt{complex, $\complex$}
\lft{top, $\top$}	\ctr{bot, $\bot$}	\rt{aleph, $\aleph$}
\lft{iit, $\iit$}	\ctr{jit, $\jit$}	\rt{angle, $\angle$}
\lft{infty, $\infty$}	\ctr{emptyset, $\emptyset$}}

The following commands are useful as paired delimiters:{\fill 0\ss{}

\lft{lfloor, $\lfloor$}		\rt{rfloor, $\rfloor$}
\lft{lceil, $\lceil$}		\rt{rceil, $\rceil$}
\lft{langle, $\langle$}		\rt{rangle, $\rangle$}
\lft{dleft, $\dleft$}		\rt{dright, $\dright$}}

There are many operators that appear in formulas that {\it should
not} be set in the same face as the surrounding formula. These
are usually set in the face of the surrounding text. tex provides
the following such operators:{\fill 0\ss{}

\lft{cos, $\cos$}	\ctr{cot, $\cot$}	\rt{csc, $\csc$}
\lft{sin, $\sin$}	\ctr{sec, $\sec$}	\rt{tan, $\tan$}
\lft{exp, $\exp$}	\ctr{ln, $\ln$}		\rt{log, $\log$}
\lft{lim, $\lim$}	\ctr{liminf, $\liminf$}	\rt{limsup, $\limsup$}
\lft{min, $\min$}	\ctr{max, $\max$}	\rt{Pr, $\Pr$}
\lft{sup, $\sup$}	\ctr{inf, $\inf$}	\rt{lg, $\lg$}
\lft{det, $\det$}	\ctr{gcd, $\gcd$}	\rt{modop, $\modop$}}

There are three math mode commands that use arguments. These are
the super and sub script command, and the modulo operator. The
following table is similar to the preceding tables, except that
the string used to generate the glyph is {\ss A\\?\{exg\},} where
the {\ss ?} is the command entered in the table. {\fill 0\ss{}

\lft{mod, $A\mod{exg}$}	\ctr{+, $A\+{exg}$}	\rt{-, $A\-{exg}$}}

Finally, there is a collection of math mode commands to generate
miscellaneous characters that are not associated with
mathematics, but are generally useful. These are listed in the
section on math mode, and repeated here for completeness.{\fill 0\ss{}

\lft{$\$$, $\$$}	\ctr{$\#$, $\#$}	\rt{dag, $\dag$}
\lft{ddag, $\ddag$}	\ctr{ldots, $\ldots$}	\rt{ldotss, $\ldotss$}
\lft{circ, $\circ$}	\ctr{bullet, $\bullet$}	\rt{copyright, $\copyright$}
\lft{section, $\section$}}

\ctr{max, $\max$}